### 2016-12-07 EC17-submitted/notions.tex
EC17-submitted
pass...

\section{Security Notions}
\label{sec:security-notions}
Following NY, we are interested in defining an adaptive (i.e.,adversarial) notion of correctness.
In addition to correctness, we establish two kinds of privacy notions for data structures.

@@

\begin{definition}
A set-membership data structure $(\Rep, \Qry)$ is {\sf canonical} if
there exist algorithms $\Rep_1, \Rep_2$, a query-enumeration
function~$\ell\colon\mathbb{N}\to\mathbb{N}$,
and a function family
$F: \calK \times \bits^{\ell_{in}} \rightarrow \bits^{\ell_{out}}$
such that the following hold:
\begin{itemize}
\item $\Rep_1$ takes as input a set $S$ and outputs distinct values $s_1, \ldots, s_t \in \bits^{\ell_{in}}$,
where $t=\ell(|S|)$
\item $\Rep_2$ takes as input $|S|$ and $y_1, \ldots, y_t \in \bits^{\ell_{out}}$, and outputs
$\pubaux$.
\item For any $S$, the distribution on $\pubaux$ output by $\Rep(S)$ is identical to the distribution
on $\pubaux$ computing as follows:
(1) $s_1, \ldots, s_t \getsr \Rep_1(S)$;
(2) $K \getsr \calK$;
(3) $\pubaux \getsr \Rep_2(|S|, F_{K}(s_1), \ldots,F_{K}(s_t))$.
%\item $K_1,K_2,\ldots,K_p \getsr \calK$ for some fixed number of keys~$p$;
%\item $\pubaux \getsr \Rep_2(n, \{F_{K_j}(s_1),
%\ldots,F_{K_j}(s_t)\}_{j\in[p]})$.
\end{itemize}
We call $(\Rep_1, \Rep_2, \ell, F)$ the {\sf canonical representation}
of algorithm~$\Rep$. \hfill\dqed
\end{definition}

>>>>

\section{Security Notions}
\label{sec:security-notions}
Following NY, we are interested in defining an adaptive (i.e., adversarial) notion of correctness.
In addition to correctness, we establish two kinds of privacy notions for data structures.

@@

\begin{definition}
A set-membership data structure $(\Rep, \Qry)$ is {\sf canonical} if
there exist algorithms $\Rep_1, \Rep_2$, a query-enumeration
function~$\ell\colon\mathbb{N}\to\mathbb{N}$,
and a function family
$F: \calK \times \bits^{\ell_{in}} \rightarrow \bits^{\ell_{out}}$
such that the following hold:
\begin{itemize}
\item $\Rep_1$ takes as input a set $S$ and outputs distinct values $s_1, \ldots, s_t \in \bits^{\ell_{in}}$,
where $t=\ell(|S|)$
\item $\Rep_2$ takes as input $|S|$ and $y_1, \ldots, y_t \in \bits^{\ell_{out}}$, and outputs
$\pubaux$.
\item For any $S$, the distribution on $\pubaux$ output by $\Rep(S)$ is identical to the distribution
on $\pubaux$ computed as follows:
(1) $s_1, \ldots, s_t \getsr \Rep_1(S)$;
(2) $K \getsr \calK$;
(3) $\pubaux \getsr \Rep_2(|S|, F_{K}(s_1), \ldots,F_{K}(s_t))$.
%\item $K_1,K_2,\ldots,K_p \getsr \calK$ for some fixed number of keys~$p$;
%\item $\pubaux \getsr \Rep_2(n, \{F_{K_j}(s_1),
%\ldots,F_{K_j}(s_t)\}_{j\in[p]})$.
\end{itemize}
We call $(\Rep_1, \Rep_2, \ell, F)$ the {\sf canonical representation}
of algorithm~$\Rep$. \hfill\dqed
\end{definition}

### 2016-12-07 abstract.tex
typos
pass...

We initiate a provable-security treatment of abstract data structures, by formalizing them as syntactic primitives, and considering their correctness and privacy first-class security properties.  Our syntax is quite broad, capturing a wide variety of in-use and academic data structures.  Concretely, we explore the security of several data structures that support (approximate) set-membership tests.  Our notion of correctness captures an adversaries ability to cause the data structure to err via its allowed set of queries.  Our main privacy notion neatly captures what is leaked by the public portion of the data structure.  As certain widely used set-membership data structures, like the Bloom filter, do not meet our notion, we consider a secondary notion of onewayness; essentially, capturing the inability to reproduce any element of the set given the public portion of the data structure.  We also compare our notions to recent work by Naor and Yogev (Crypto'15), which also considers Bloom-filter-like data structures, albeit in a much more limited way.  Our syntax and notions let us consider security under a variety of usage scenarios, by clearly surfacing the split between public and private portions of the data structure.

We see this work as laying the foundation for a number of follow-on efforts, for example to study dynamic data structures (e.g. Bloom filter variants whose representation changes over time), and the security of primitives built on top of abstract data structures.

>>>>

We initiate a provable-security treatment of abstract data
structures by formalizing them as syntactic primitives and
considering their correctness and privacy as first-class security
properties.  Concretely, we explore the security of several data
structures that support (approximate) set-membership tests.  Our
notion of correctness captures an adversary's inability to cause a
data structure to err via its allowed set of queries, and our
privacy notion neatly captures what is leaked by the public portion
of a data structure.  As certain widely used set-membership data
structures, such as Bloom filters, do not meet our privacy notion,
we consider a secondary notion of onewayness capturing the inability
to recover any element of a high-entropy set given the public
portion of the data structure. We also compare our notions to recent
work by Naor and Yogev (Crypto 2015), which also considers
Bloom-filter-like data structures, albeit in a much more limited
way.  Our syntax and notions let us consider security under a
variety of usage scenarios, by clearly surfacing the split between
public and private portions of the data structure.

%We see this work as laying the foundation for a number of follow-on
%efforts, for example to study dynamic data structures (e.g., Bloom
%filter variants whose representation changes over time), and the
%security of primitives built on top of abstract data structures.

### 2016-12-07 intro.tex
pass...

Data structures are a fundamental tool used throughout computerscience. Typically, when the behavior of a data structure (e.g., its running time or its error probability) is analyzed, the analysis is performed under the assumption that any inputs and queries to the data structure are chosen independently of the data structure itself.
In particular, independent of any internal randomness used by the data structure. But as recently highlighted by Naor and Yogev~\cite{naor2015bloom}, there are many natural scenarios in which the input and queries may be chosen \emph{adversarially} and \emph{adaptively}, based on partial information and prior observations about the data structure.

In addition, standard applications of data structures assume that the party querying the data structure is fully trusted. Hence, privacy concerns do not arise. But in many real-world scenarios, a data structure is built from sensitive data.  Here one would like to know what information is leaked by queries to the data structure; minimally, to know that a reasonable number of queries to not effectively divulge the entire data.  Existing definitions say nothing about privacy in this sense.

These concerns motivate a careful and formal treatment of data structures in an adversarial setting, 
where both \emph{correctness} and \emph{privacy} are considered. In this work we initiate such a
treatment.

\heading{Formalizing abstract data structures.} Loosely speaking, we formalize a data structure as a pair of algorithms $(\Rep,\Qry)$, the representation and query-evaluation algorithms, respectively.  The randomized representation algorithm takes in an abstract object~$S$ and produces public representation data $\pubaux$ and private representation data $\privaux$.  Intuitively, $\pubaux$ is considered public information, whereas $\privaux$ is meant to be known only to the party running $\Qry$.  The deterministic query-evaluation algorithm uses $\pubaux$ and $\privaux$ in order to (approximately) evaluate permitted queries $\qry$, which are arbitrary functions over abstract objects.  We say ``approximately''  evaluate permitted queries because $\pubaux$ and $\privaux$ may not contain all of the information sufficient to determine the true value of $\qry(S)$ for all~$S$.

Let us give an example.  Consider the classic Bloom filter~\cite{bloom1970space}, in which case $\pubaux$ will consist of the bit array~$M$ that represents a given set~$S$, as well as descriptions of the hash functions $H_1,H_2,\ldots,H_k$ used to produce~$M$; $\privaux$ will be empty.  The permitted queries will be predicates of the form $\qry_x(S)=1$ iff $x \in S$. The~$\Qry$ algorithm, on input $M,H_1,H_2,\ldots,H_k$ and $\qry_x$, carries out the standard Bloom filter set-membership test to decide whether or not $x \in S$.  It is well known that Bloom fiters admit false-positive errors, thus $\Qry$ only approximates the true set-membership test.  Our correctness security notion will quantitatively capture the approximation error, but in the presense of attackers that adaptively try to induce one or more errors (false-positives, in this case).

By shifting $H_1,H_2,\ldots,H_k$ from $\pubaux$ to $\privaux$, we do not change the functionality at all.  But we may very well change the security profile of the resulting Bloom filter variant.  Indeed, this is the case.  When $H_1,H_2,\ldots,H_K$ are assumed to be good pseudorandom functions, we can prove that this variant achieves a kind of ``semantic security'' privacy guarantee: the bit array~$M$ is well known to leak a good approximation of~$|S|$, and our notion shows that it leaks nothing more than this.  (Such a result is not possible for the standard Bloom filter, or any other data structure with empty~$\privaux$.)

\heading{Case Studies.}
This example offers a hint at the broad range of abstract data structures our syntax supports.  In this paper, we put the syntax and our security notions to work in several case studies.  First, we analyze classical Bloom filters when the hash functions are modeled as public random oracles~\cite{broder2004network}, when the hash functions are secretly-keyed PRFs ~\cite{niedermeyer2014cryptanalysis,naor2015bloom}, and when the hash functions are secretly keyed but fall short of being PRFs, i.e., so-called ``linear-PRF'' hash functions~\cite{kirsch2006less,niedermeyer2014cryptanalysis}.  Each of these represents an in-use or proposed Bloom filter variant.

The linear-PRF Bloom filter provides an opportunity to see just how brittle constructions can be to their usage.  When used to create a representation of a set, we prove that the linear-PRF Bloom filter offers both correctness and privacy.  However, when applied to (effectively) a multiset privacy is lost.  In fact, Niedermeyer et al.~\cite{niedermeyer2014cryptanalysis} show that the linear-PRF Bloom filter, when applied to a collection of bigrams of German last names, admits a practical attack that reveals the names.  Note that this is despite the fact that the hash functions are constructed from secretly-keyed PRFs in a way that has been suggested in recent literature on Bloom filter variants~\cite{kirsch2006less}.

We also consider garbled Bloom filters, proposed by Dong et al.~\cite{dong2013private} as a key piece of a private set-intersection protocol.

We see this work as laying the foundation for a number of follow-on studies.  For example to study dynamic data structures, such as counting Bloom filters\cite{fan2000summary}, scalable Bloom filters~\cite{almeida2007scalable}, count-min sketches~\cite{cormode2005improved}, hierachical Bloom filters~\cite{zhu2004hierarchical}, to name just a few in the extended Bloom filter family.  Another direction would be to explore the security of primitives built \emph{from} these sorts of data structures, such as content-distribution networks (where many servers propogate representations of their local cache to their neighbors), or cryptographic strike lists (a la early versions of the QUIC protocol) for nonce-replay protection.  The Bloom filter family alone has a wide range of practical applications, for example in large database query
processing, routing algorithms for peer-to-peer networks, protocols
for establishing linkages between medical-record databases, fair
routing of TCP packets, and Bitcoin wallet synchonization~\cite{schnell2011novel,niedermeyer2014cryptanalysis,gervais2014privacy,nojima2009cryptographically,feng2001stochastic,reynolds2003efficient,byers2002informed,broder2004network}.

\heading{Comparison to~\cite{naor2015bloom}.} Naor and
Yogev~\cite{naor2015bloom} recently considered the problem of
adversarial soundness for Bloom filters and, indeed,
their work served as inspiration for our own. Our work extends
their work significantly in several directions. For starters, we
consider abstract data structures (rather than only the case of
set-membership data structures), and consider privacy in addition to
correctness. Even with respect to the specific case of correctness for
set-membership data structures, our work offers several advantages
to the Naor-Yogev treatment:
\begin{itemize}
\item We explicitly distinguish between public and private
    portions of a data structure, something that was left
    ambiguous\footnote{In fact, it isn't clear that their formalization actually allows for standard Bloom filters, although fixing this ambiguity is straightforward once recognized.} in~\cite{naor2015bloom}. This allows us, for
    example, to distinguish between the hash-function keys and
    the bit array when discussing classical Bloom filters.
\item The Naor-Yogev definition of soundness allows the
    adversary to make several set-membership queries, some of
    which may produce incorrect results, and the attacker then
    succeeds if it can output a \emph{fresh} query which leads
    to an error. We propose instead what we find to be a more
    natural definition in which the attacker succeeds if it can
    cause a certain number of incorrect results.
\item Naor and Yogev analyze a new Bloom-filter construction of
    their own design. In contrast, we are interested in
    analyzing practical constructions that have been proposed in
    prior work to understand how secure they are. In particular,
    we analyze Bloom filters in the random-oracle model.
\end{itemize}

We give a technical comparison of our correctness notion and theirs (lifted to our more general syntax)
in Appendix~\ref{sec:compare-defs}.

>>>>

Data structures are a fundamental tool used throughout computer
science. Typically, when the behavior of a data structure (e.g., its
running time or its error probability) is analyzed, the analysis is
performed under the assumption that any inputs and queries to the
data structure are chosen non-adversarially and, in particular, are
independent of any internal randomness used by the data structure.
But as recently highlighted by Naor and Yogev~\cite{C:NaoYog15},
there are many natural scenarios where the input and/or queries may
be chosen \emph{adversarially} and \emph{adaptively}, based on
partial information and prior observations about the data structure.

In addition to the above, standard applications of data structures
assume that the party querying the data structure is fully trusted
and hence privacy concerns do not arise. But in many real-world
scenarios, a data structure is built from sensitive data.  Here, it
is important to understand what information is leaked by public
portions of the data structure and, minimally, to know that a
reasonable number of queries will not divulge ``too much more
information'' that was is implied by the answers to those queries.
Existing definitions say nothing about privacy in this sense.

These concerns motivate a careful and formal treatment of data
structures in an adversarial setting, where both \emph{correctness}
and \emph{privacy} are considered. We initiate such a treatment in
this work.

\heading{Formalizing abstract data structures.} Loosely speaking, we
formalize a data structure as a pair of algorithms $(\Rep,\Qry)$,
the representation and query-evaluation algorithms, respectively.
The randomized representation algorithm takes in an abstract
object~$S$ and produces public representation data $\pubaux$ and
private representation data $\privaux$.  Intuitively, $\pubaux$ is
public information, whereas $\privaux$ is meant to be known only to
the party running $\Qry$.  The deterministic query-evaluation
algorithm uses $\pubaux$ and $\privaux$ in order to (approximately)
evaluate a supported query $\qry$, which can in principle be an
arbitrary function.
%We say ``approximately'' evaluate permitted queries
%because $\pubaux$ and $\privaux$ may not contain all of the
%information sufficient to determine the true value of $\qry(S)$ for
%all~$S$.

As an example, consider the classic Bloom
filter~\cite{bloom1970space}. Here, $\pubaux$ consists of a bit
array~$M$ that represents a given set~$S$, as well as descriptions
of the hash functions $H_1,H_2,\ldots,H_k$ used to produce~$M$;
$\privaux$ is empty.  The supported queries are predicates of the
form $\qry_x(S)=1$ iff $x \in S$. The~$\Qry$ algorithm, on input
$M,H_1,\ldots,H_k$ and $\qry_x$, carries out the standard
Bloom-filter test to determine if $x \in S$. It is well known that
Bloom filters admit false-positive errors; thus, $\Qry$ only
approximates the true set-membership test.  Our correctness notion
will quantitatively capture the error, even in the presence of
attackers that adaptively try to induce one or more errors
(false-positives, in this case).

By shifting $H_1,H_2,\ldots,H_k$ from $\pubaux$ to $\privaux$, we do
not change the functionality, but we may very well change the
security profile of the resulting Bloom-filter variant.  Indeed,
this is the case.  When $H_1,H_2,\ldots,H_K$ are assumed to be
pseudorandom functions, we can prove that this variant achieves a
sort of ``semantic security'' privacy guarantee: the bit array~$M$
leaks a good approximation of~$|S|$, but nothing more.  (Such a
result is not possible for the standard Bloom filter, or any other
data structure with empty~$\privaux$.)

\heading{Case studies.} The example above offers a hint at the broad
range of abstract data structures our syntax supports.  In this
paper, we put the syntax and our security notions to work in several
case studies.  First, we analyze classical Bloom filters when the
hash functions are modeled as public random
oracles~\cite{broder2004network}, when the hash functions are
secretly-keyed PRFs~\cite{niedermeyer2014cryptanalysis,C:NaoYog15},
and when the hash functions are secretly keyed but fall short of
being PRFs, i.e., so-called ``linear-PRF'' hash
functions~\cite{kirsch2006less,niedermeyer2014cryptanalysis}.  Each
of these represents an in-use or proposed Bloom-filter variant. We
also consider garbled Bloom filters~\cite{CCS:DonCheWen13}, proposed
as a key piece of a private set-intersection protocol.

The linear-PRF Bloom filter illustrates how brittle constructions
can be depending on how they are used. When used to create a
representation of a \emph{set}, we prove that the linear-PRF Bloom
filter offers both correctness and privacy. However, when applied to
a \emph{multiset}, privacy is lost.  In fact, Niedermeyer et
al.~\cite{niedermeyer2014cryptanalysis} show that the linear-PRF
Bloom filter, when applied to a collection of bigrams derived from
German last names, admits a practical attack that reveals the names,
despite the fact that the hash functions are constructed from
secretly-keyed PRFs in a way that has been suggested in recent
literature on Bloom-filter variants~\cite{kirsch2006less}.

%We see our work as laying the foundation for a number of follow-on
%studies.  For example to study dynamic data structures, such as
%counting Bloom filters\cite{fan2000summary}, scalable Bloom
%filters~\cite{almeida2007scalable}, count-min
%sketches~\cite{cormode2005improved}, hierarchical Bloom
%filters~\cite{zhu2004hierarchical}, to name just a few in the
%extended Bloom filter family.  Another direction would be to explore
%the security of primitives built \emph{from} these sorts of data
%structures, such as content-distribution networks (where many
%servers propagate representations of their local cache to their
%neighbors), or cryptographic strike lists (a la early versions of
%the QUIC protocol) for nonce-replay protection.  The Bloom filter
%family alone has a wide range of practical applications, for example
%in large database query processing, routing algorithms for
%peer-to-peer networks, protocols for establishing linkages between
%medical-record databases, fair routing of TCP packets, and Bitcoin
%wallet
%synchronization~\cite{schnell2011novel,niedermeyer2014cryptanalysis,gervais2014privacy,nojima2009cryptographically,feng2001stochastic,reynolds2003efficient,byers2002informed,broder2004network}.

### 2016-12-07 related-work.tex
pass...

The Naor-Yogev paper~\cite{naor15bloom} served as inspiration for our work.  We recommend reading the very good related work section contained in their paper for a discussion of related theory papers.  Here we mention a few additional practical works, but stress that this only scratches the surface.

As previously mentioned Bloom filters and their relatives are some of the most widely used data structures supporting set-membership queries.  We give a few more, here. Hbase, a hadoop based NoSql database designed to handle large datasets has an implementation of Bloom filter and counting Bloom filter. Hbase is the open-source implementation of Google's BigTable storage system~\cite{chang2008bigtable}. 
The Squid proxy~\cite{fan2000summary} uses a bloom filter as a ``summary cache'' of the set of URLs in its cache in order to web object retreival latency.   Crossby and Walach~\cite{crosby2003denial} exploited hash collisions in Squid to increase the average URL load time.  In an older paper, Lipton and Naughton~\cite{lipton1993clocked} showed how timing analysis of record insertion in a hash table can help an adversary to adapt and chose elements that take more time to be mapped into the hash table. Reynolds and Vahdat~\cite{reynolds2003efficient} proposed an efficient distributed search engine that can be used to search for files that contain a particular keyword. This search engine maps the keywords of each file into a Bloom filter; a look-up of the keyword in the Bloom filter tells whether the node has files containing that keyword or not. Record linkage is very often required to collect and combine records from different databases.

Schnell et. al.~\cite{schnell2011novel} proposed the so-called ``Cryptographic Longterm Key'' as a way to support privacy-preserving record linkage. This uses secretly keyed hash functions to create a Bloom filter for each surname. The surname is broken into bigrams and mapped into the Bloom filter with secret hash functions.  Subsequently, Niedermeyer et al.~\cite{niedermeyer2014cryptanalysis} showed an attack on a database of German surnames which were stored in secretly hashed Bloom filters, similar to Cryptographic Longterm Key.

Stochastic Fair Blue~\cite{feng2001stochastic} uses counting Bloom filter to manage non-responsive TCP traffic.

>>>>

Naor and Yogev~\cite{C:NaoYog15} recently considered the problem of
adversarial soundness for Bloom filters and, indeed, their work
served as inspiration for our own. Our work extends their work
significantly in several directions. For starters, we consider
abstract data structures (rather than only the case of
set-membership data structures), and consider privacy in addition to
correctness. Even with respect to the specific case of correctness
for set-membership data structures, our work offers several
advantages to the Naor-Yogev treatment:
\begin{itemize}
\item We explicitly distinguish between public and private
    portions of a data structure, something they leave
    ambiguous.\footnote{In fact, it isn't clear that their
formalization actually allows for standard Bloom filters,
    although fixing this ambiguity is straightforward once
    recognized.} This allows us, for example, to distinguish
    between the hash-function keys and the bit array when
    discussing classical Bloom filters.
\item The Naor-Yogev definition of soundness allows the
    adversary to make several set-membership queries, some of
    which may produce incorrect results, and the attacker then
    succeeds if it can output a \emph{fresh} query which leads
    to an error. We propose instead what we find to be a more
    natural definition in which the attacker succeeds if it can
    cause a certain number of incorrect results.
\item Naor and Yogev analyze a new Bloom-filter construction of
    their own design. In contrast, we are interested in
    analyzing practical constructions that have been proposed in
    prior work to understand their security. In particular, we
    analyze Bloom filters in the random-oracle model, something
    not considered by Naor and Yogev.
\end{itemize}

%We give a technical comparison of our correctness notion and theirs (lifted to our more general syntax)
%in Appendix~\ref{sec:compare-defs}.

We recommend reading the Naor-Yogev paper for a survey of related
work and a discussion of related theory papers. Here we mention a
few additional practical works, but stress that this only scratches
the surface.

\jnote{Rather random collection of papers using Bloom filters and
variants. I removed it for now, since it's not clear that they have
any particular relevance to us.}
%As previously mentioned, Bloom
%filters and their relatives are some of the most widely used data
%structures supporting set-membership queries. As examples, Hbase,
%the open-source implementation of Google's BigTable storage
%system~\cite{chang2008bigtable},
%%a Hadoop-based NoSql database designed to handle large datasets,
%includes an implementation of Bloom filters and counting Bloom
%filters, and the Squid proxy~\cite{fan2000summary} uses a Bloom
%filter as a ``summary'' of the set of URLs in its cache in order to
%improve latency for web-object retrieval. Reynolds and
%Vahdat~\cite{reynolds2003efficient} proposed an efficient
%distributed search engine that can be used to search for files
%containing a particular keyword. Their search engine maps the
%keywords of each file into a Bloom filter; a look-up of the keyword
%in the Bloom filter tells whether the node has files containing that
%keyword or not. Stochastic Fair Blue~\cite{feng2001stochastic} uses
%counting Bloom filter to manage non-responsive TCP traffic.

Lipton and Naughton~\cite{lipton1993clocked} showed how timing
analysis of record insertion in a hash table can help an adversary
to adaptively chose elements so as to increase the running time.
Crosby and Walach~\cite{crosby2003denial} exploited hash collisions
in Squid to increase the average URL load time.

Schnell et al.~\cite{schnell2011novel} proposed a way to support
privacy-preserving record linkage by using secretly keyed hash
functions. Subsequently, Niedermeyer et
al.~\cite{niedermeyer2014cryptanalysis} showed an attack on a
similar system when used on a database of German surnames.
\jnote{See Section~???}
